---
title: "Jackson Heart Study Data Analysis"
output: html_document
date: "2024-03-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The Jackson Heart Study is a community-based cohort study evaluating the etiology of cardiovascular diseases among African Americans residing in Hinds, Madison, and Rankin counties in the Jackson, Mississippi metropolitan area. Data was collected from 5306 participants and we will be conducing an analysis on an existing Jackson Heart Study data set.

The research question for this analysis is:

***Is smoking associated with hypertension after controlling for socio-demographic variables?***

```{r}
library(here)
library(rprojroot)

rprojroot::find_rstudio_root_file()

# Move data set into project directory data folder
file.copy(from = "/Users/drendon/Desktop/baseline_jhs.csv",
          to   = "/Users/drendon/BIODS253 Final/Data/baseline_jhs.csv")
```

## Data Exploration

First, we will explore the data structure and variables.

```{r}
library(tidyverse)
jhs_data = read_csv(here("data/baseline_jhs.csv"))
head(jhs_data)

```

-   Each column represents a variable and each row represents a patient.

    ```{r}
    #check continuous variables - age
    hist(jhs_data$ageV1, breaks = 100)
    summary(jhs_data$ageV1)

    #check continuous variables - alcohol
    hist(jhs_data$alcwV1)

    ```

-   Age variable

    -   Results show patient with age = 999

```{r}
# check categorical variables
table(jhs_data$maleV1, useNA = "ifany")
table(jhs_data$IncomeV1, useNA = "ifany")
table(jhs_data$edu3catV1, useNA = "ifany")

#check missing percentages for income based on check above
mean(is.na(jhs_data$IncomeV1))

```

## Data Cleaning

Based on some of the findings above, we will conduct some data cleaning.

```{r}
#Assume anyone over 125 years does not have a valid age and create a new dataset
clean_jhs_data = jhs_data %>% 
  mutate(ageV1 = if_else(ageV1>120, NA, ageV1))
```
