```{r}
library(tidyverse)
library(here)
library(ggplot2)
```

## Individual data

```{r}
individual_data = read_csv(here("Final Project: Dengue Risk Factor Analysis","Data","Raw Data","dengue_individual_data.csv"))
glimpse(individual_data)

```

**Data dictionary:**

| Variable Name        | Description                                                                              |
|------------------|------------------------------------------------------|
| `ID`                 | Individual identifier                                                                    |
| `hh_ID`              | Household identifier                                                                     |
| `sex`                | Biological sex                                                                           |
| `n_symp_infections`  | Count of times this individual has had a symptomatic dengue infection during the study   |
| `n_asymp_infections` | Count of times this individual has had an asymptomatic dengue infection during the study |

## Household data

```{r}
household_data = read_csv(here("Final Project: Dengue Risk Factor Analysis","Data","Raw Data","dengue_household_data.csv"))
glimpse(household_data)
```

**Data Dictionary:**

| Variable Name           | Description                                                                                                                                                                                                                                                                                                       |
|----------------|--------------------------------------------------------|
| `hh_ID`                 | Household identifier                                                                                                                                                                                                                                                                                              |
| `door_screens`          | 1, if the household has a screen door. 0, otherwise                                                                                                                                                                                                                                                               |
| `house_type`            | One of three household types. "Poles" indicates the house is built on stilts, as homes on floodplains typically are. "Single" indicates the house is a single-family home separate from other homes. "Townhouse" indicates the house if part of a multifamily unit where multiple homes are physically connected. |
| `concrete_house`        | 1, if the house was built from concrete. 0, otherwise                                                                                                                                                                                                                                                             |
| `w ater_supply_by_pipe` | 1, if the house is supplied water by pipes. 0, if the house receives water some other way                                                                                                                                                                                                                         |
| `garbage_management`    | One of three methods for getting rid of household waste. "Burn" if it is burned by the residents. "Car" if it is disposed of by a vehicle that collects waste. "Dump" if it is brought to a waste disposal facility by the residents.                                                                             |
| `n_toilets_outside`     | Count of the outdoor toilets available to the household residents                                                                                                                                                                                                                                                 |

# Data Exploration

```{r}
#see part of the data
names(individual_data)
names(household_data)

head(individual_data)
head(household_data)

#explore distribution of continuous variable age in the individual dataset
hist(individual_data$age)
summary(individual_data$age)

#explore distribution of the categorical variables n_symp_infections and n_asymp_infections in the individual dataset
table(individual_data$n_symp_infections)
table(individual_data$n_asymp_infections)
table(individual_data$sex)

#explore distribution of the categorical variables door_screens, concrete_house, water_supply_by_pipe, and n_toilets_outside in the individual dataset
table(household_data$door_screens)
table(household_data$concrete_house)
table(household_data$water_supply_by_pipe)
table(household_data$n_toilets_outside)

#Ensure that there are no missing hh_IDs from both datasets
summary(individual_data$hh_ID)
summary(household_data$hh_ID)

#check for missing data
prop.table(table(is.na(individual_data$age)))
prop.table(table(is.na(individual_data$sex)))
prop.table(table(is.na(individual_data$n_symp_infections)))
prop.table(table(is.na(individual_data$n_asymp_infections)))

prop.table(table(is.na(household_data$door_screens)))
prop.table(table(is.na(household_data$concrete_house)))
prop.table(table(is.na(household_data$water_supply_by_pipe)))
prop.table(table(is.na(household_data$n_toilets_outside)))
```

-   There are missing values in the water_supply_by_pipe variable in the household_data

-   Missing values are coded as "Unknown"

```{r}
#find the percentage of missing values for the water_supply_by_pipe variable in household_data
total_water_supply_by_pipe_count = household_data %>% 
  nrow()

missing_water_supply_by_pipe_count = sum(household_data$water_supply_by_pipe == "Unknown")

missing_water_supply_by_pipe_count

percent_missing_water_supply_by_pipe = (missing_water_supply_by_pipe_count / total_water_supply_by_pipe_count) *100

percent_missing_water_supply_by_pipe
```

-   There are 95 missing/unknown values for water_supply_by_pipe
-   About 10% of values are missing for that variable
