```{r}
library(tidyverse)
library(here)
library(ggplot2)
```

## Individual data

```{r}
individual_data = read_csv(here("Final Project: Dengue Risk Factor Analysis","Data","Raw Data","dengue_individual_data.csv"))
glimpse(individual_data)

```

**Data dictionary:**

| Variable Name        | Description                                                                              |
|------------------|------------------------------------------------------|
| `ID`                 | Individual identifier                                                                    |
| `hh_ID`              | Household identifier                                                                     |
| `sex`                | Biological sex                                                                           |
| `n_symp_infections`  | Count of times this individual has had a symptomatic dengue infection during the study   |
| `n_asymp_infections` | Count of times this individual has had an asymptomatic dengue infection during the study |

## Household data

```{r}
household_data = read_csv(here("Final Project: Dengue Risk Factor Analysis","Data","Raw Data","dengue_household_data.csv"))
glimpse(household_data)
```

**Data Dictionary:**

| Variable Name           | Description                                                                                                                                                                                                                                                                                                       |
|----------------|--------------------------------------------------------|
| `hh_ID`                 | Household identifier                                                                                                                                                                                                                                                                                              |
| `door_screens`          | 1, if the household has a screen door. 0, otherwise                                                                                                                                                                                                                                                               |
| `house_type`            | One of three household types. "Poles" indicates the house is built on stilts, as homes on floodplains typically are. "Single" indicates the house is a single-family home separate from other homes. "Townhouse" indicates the house if part of a multifamily unit where multiple homes are physically connected. |
| `concrete_house`        | 1, if the house was built from concrete. 0, otherwise                                                                                                                                                                                                                                                             |
| `w ater_supply_by_pipe` | 1, if the house is supplied water by pipes. 0, if the house receives water some other way                                                                                                                                                                                                                         |
| `garbage_management`    | One of three methods for getting rid of household waste. "Burn" if it is burned by the residents. "Car" if it is disposed of by a vehicle that collects waste. "Dump" if it is brought to a waste disposal facility by the residents.                                                                             |
| `n_toilets_outside`     | Count of the outdoor toilets available to the household residents                                                                                                                                                                                                                                                 |

# Data Exploration

```{r}
#see part of the data
names(individual_data)
names(household_data)

head(individual_data)
head(household_data)

#explore distribution of continuous variable age in the individual dataset
hist(individual_data$age)
summary(individual_data$age)

#explore distribution of the categorical variables n_symp_infections and n_asymp_infections in the individual dataset
table(individual_data$n_symp_infections)
table(individual_data$n_asymp_infections)
table(individual_data$sex)

#explore distribution of the categorical variables door_screens, concrete_house, water_supply_by_pipe, and n_toilets_outside in the individual dataset
table(household_data$door_screens)
table(household_data$concrete_house)
table(household_data$water_supply_by_pipe)
table(household_data$n_toilets_outside)

#Ensure that there are no missing hh_IDs from both datasets
summary(individual_data$hh_ID)
summary(household_data$hh_ID)

#check for missing data
prop.table(table(is.na(individual_data$age)))
prop.table(table(is.na(individual_data$sex)))
prop.table(table(is.na(individual_data$n_symp_infections)))
prop.table(table(is.na(individual_data$n_asymp_infections)))

prop.table(table(is.na(household_data$door_screens)))
prop.table(table(is.na(household_data$concrete_house)))
prop.table(table(is.na(household_data$water_supply_by_pipe)))
prop.table(table(is.na(household_data$n_toilets_outside)))
```

-   There are missing values in the water_supply_by_pipe variable in the household_data

-   Missing values are coded as "Unknown"

```{r}
#find the percentage of missing values for the water_supply_by_pipe variable in household_data
total_water_supply_by_pipe_count = household_data %>% 
  nrow()

missing_water_supply_by_pipe_count = sum(household_data$water_supply_by_pipe == "Unknown")

missing_water_supply_by_pipe_count

percent_missing_water_supply_by_pipe = (missing_water_supply_by_pipe_count / total_water_supply_by_pipe_count) *100

percent_missing_water_supply_by_pipe
```

-   There are 95 missing/unknown values for water_supply_by_pipe
-   About 10% of values are missing for that variable

# Data Cleaning

```{r}
#create copies of individual and household data
individual_data_cleaned = individual_data
household_data_cleaned = household_data

#recode missing/unknown values in water_supply_by_pipe variable to NA and to numerics
household_data_cleaned = household_data_cleaned %>% 
  mutate(water_supply_by_pipe_clean = ifelse(water_supply_by_pipe == "Unknown", NA,
                                      as.numeric(water_supply_by_pipe)))

#check numerics
check_numeric = sapply(household_data_cleaned$water_supply_by_pipe_clean, is.numeric)
check_numeric

```

Add the following variables to individual_data_cleaned:

-   `age_cat`, that is "adult" for people 18 or older and "child" for people under 18.

-   `n_any_infections`, the sum of symptomatic and asymptomatic infections for each person.

-   `any_infections`, that is `TRUE` if the individual was ever infected with symptomatic or asymptomatic dengue.

```{r}
individual_data_cleaned = individual_data_cleaned %>% 
  mutate(age_cat = ifelse(age >= 18, "Adult", "Child"),
         n_any_infections = n_symp_infections + n_asymp_infections,
         any_infections = (n_symp_infections > 0 | n_asymp_infections > 0))

#check new variables
select(individual_data_cleaned, age, age_cat)

select(individual_data_cleaned, n_symp_infections, n_asymp_infections, n_any_infections, any_infections)

```

-   Join household and individual data sets by common household ID variable. Resulting data frame contains 1 row per individual with columns for both individual and household level variables.

```{r}
#merge individual_data_cleaned and household_data_cleaned (full_join to ensure all existing and new variables are included in the merged dataset)
merged_dengue_data = full_join(individual_data_cleaned, household_data_cleaned, by = "hh_ID")

#check merged dataset
head(merged_dengue_data)

```

```{r}
#save new data sets into processed data folder
write.csv(individual_data_cleaned, file = here("Final Project: Dengue Risk Factor Analysis/Data/Processed Data/individual_data_cleaned.csv"), row.names = FALSE)

write.csv(household_data_cleaned, file = here("Final Project: Dengue Risk Factor Analysis/Data/Processed Data/household_data_cleaned.csv"), row.names = FALSE)


```
