```{r}
#import merged dengue data set to complete data analysis
merged_dengue_data = read_csv(here("Final Project: Dengue Risk Factor Analysis","Data","Processed Data","merged_dengue_data.csv"))
glimpse(merged_dengue_data)
```

```{r}
#total individuals
total_individuals = nrow(merged_dengue_data)
print(paste("Total individuals:", total_individuals))

#individuals with no infection
total_individuals_no_infection = sum(merged_dengue_data$any_infections == FALSE)
print(paste("Total individuals with no history of infection:", total_individuals_no_infection))

#individuals with any infection
total_individuals_any_infection = sum(merged_dengue_data$any_infections == TRUE)
print(paste("Total individuals with any history of infection:", total_individuals_any_infection))

```

-   Calculate age statistics stratified across infection status

```{r}
#mean age (and standard deviation) for total individuals
merged_dengue_data %>% 
  summarise(mean_age = round(mean(age), 1),
            age_SD = round(sd(age), 1))

#mean age (and standard deviation) grouped by infection history
merged_dengue_data %>% 
  group_by(any_infections) %>% 
  summarise(age_infection = round(mean(age), 1),
            age_infection_SD = round(sd(age), 1))
```

-   Calculate gender statistics stratified across infection status

```{r}
#total number of females/male and percentages
total_females = sum(merged_dengue_data$sex == "F")
print(paste("Total females:", total_females))

total_females_percent = round((total_females/total_individuals) * 100, 1)
print(paste("Total percentage of females:", total_females_percent))


total_males = sum(merged_dengue_data$sex == "M")
print(paste("Total males:", total_males))

total_males_percent = round((total_males/total_individuals) * 100, 1)
print(paste("Total percentage of males:", total_males_percent))

#Gender/infection table
gender_infection_table = table(merged_dengue_data$sex,
                               merged_dengue_data$any_infections)
gender_infection_table

gender_infection_table_percent = round(prop.table(gender_infection_table, margin = 2) * 100, 2)
gender_infection_table_percent
```

-   Calculate statistics between infection and concrete house types

    ```{r}
    #concrete and nonconcrete houses with and without infections totals
    total_concrete_houses = sum(merged_dengue_data$concrete_house == 1)
    print(paste("Total concrete houses:", total_concrete_houses))

    total_concrete_houses_percent = round((total_concrete_houses/total_individuals) * 100, 1)
    print(paste("Total percentage of concrete houses:", total_concrete_houses_percent))


    total_non_concrete_houses = sum(merged_dengue_data$concrete_house == 0)
    print(paste("Total non-concrete houses:", total_non_concrete_houses))

    total_non_concrete_houses_percent = round((total_non_concrete_houses/total_individuals) * 100, 1)
    print(paste("Total percentage of non-concrete houses:", total_non_concrete_houses_percent))

    #concrete and nonconcrete houses with and without infections
    concrete_house_infection_table = table(merged_dengue_data$concrete_house,
                                   merged_dengue_data$any_infections)
    concrete_house_infection_table

    concrete_house_infection_table_percent = round(prop.table(concrete_house_infection_table, margin = 2) * 100, 2)
    concrete_house_infection_table_percent
    ```

<!-- -->

-   Calculate statistics between infection and concrete house types

    ```{r}
    #house types: One of three household types. "Poles" indicates the house is built on stilts, as homes on floodplains typically are. "Single" indicates the house is a single-family home separate from other homes. "Townhouse" indicates the house if part of a multifamily unit where multiple homes are physically connected.

    #pole houses
    total_pole_houses = sum(merged_dengue_data$house_type == "Poles")
    print(paste("Total pole houses:", total_pole_houses))

    total_pole_houses_percent = round((total_pole_houses/total_individuals) * 100, 1)
    print(paste("Total percentage of pole houses:", total_pole_houses_percent))

    #single family houses
    total_single_houses = sum(merged_dengue_data$house_type == "Single")
    print(paste("Total single houses:", total_single_houses))

    total_single_houses_percent = round((total_single_houses/total_individuals) * 100, 1)
    print(paste("Total percentage of single houses:", total_single_houses_percent))

    #single townhouses
    total_townhouses = sum(merged_dengue_data$house_type == "Townhouse")
    print(paste("Total townhouses:", total_townhouses))

    total_townhouses_percent = round((total_townhouses/total_individuals) * 100, 1)
    print(paste("Total percentage of townhouses:", total_townhouses_percent))

    #house types with and without infections
    house_type_infection_table = table(merged_dengue_data$house_type,
                                   merged_dengue_data$any_infections)
    house_type_infection_table

    house_type_infection_table_percent = round(prop.table(house_type_infection_table, margin = 2) * 100, 2)
    house_type_infection_table_percent
    ```

-   Calculate statistics between infection and concrete house types

    ```{r}
    # 1, if the house is supplied water by pipes. 0, if the house receives water some other way
    #water supplied by pipes
    total_pipe_houses = sum(merged_dengue_data$water_supply_by_pipe_clean == 1, na.rm = TRUE)
    print(paste("Total pipe houses:", total_pipe_houses))

    total_pipe_houses_percent = round((total_pipe_houses/total_individuals) * 100, 1)
    print(paste("Total percentage of pipe houses:", total_pipe_houses_percent))

    #water not supplied by pipe
    total_non_pipe_houses = sum(merged_dengue_data$water_supply_by_pipe_clean == 0, na.rm= TRUE)
    print(paste("Total non_pipe houses:", total_non_pipe_houses))

    total_non_pipe_houses_percent = round((total_non_pipe_houses/total_individuals) * 100, 1)
    print(paste("Total percentage of non-pipe houses:", total_non_pipe_houses_percent))

    #total houses with pipe supply data missing
    total_pipe_data_missing = sum(is.na(merged_dengue_data$water_supply_by_pipe_clean))
    print(paste("Total pipe supply data missing:", total_pipe_data_missing))

    total_pipe_data_missing_percent = round((total_pipe_data_missing/total_individuals) * 100, 1)
    print(paste("Total percentage of non-pipe houses:", total_pipe_data_missing_percent))

    #house types with and without infections
    pipe_infection_table = table(merged_dengue_data$water_supply_by_pipe_clean,
                                   merged_dengue_data$any_infections,  useNA = "ifany")
    pipe_infection_table

    pipe_infection_table_percent = round(prop.table(pipe_infection_table, margin = 2) * 100, 2)
    pipe_infection_table_percent
    ```

+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Table 1                  | Total                                                                    | No Infection                                                             | Any Infection                                                            |
+==========================+:========================================================================:+:========================================================================:+:========================================================================:+
| Number of individuals, N | 4490                                                                     | 3529                                                                     | 961                                                                      |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Age, mean (SD)           | 29.2 (22.1)                                                              | 30.9 (22.3)                                                              | 22.7 (20.4)                                                              |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Sex, N (%)               | ------------------------------------------------------------------------ | ------------------------------------------------------------------------ | ------------------------------------------------------------------------ |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| \- Female                | 2599 (57.9)                                                              | 2063 (58.5)                                                              | 536 (55.8)                                                               |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| \- Male                  | 1891 (42.1)                                                              | 1466 (41.5)                                                              | 425 (44.2)                                                               |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Concrete House, N (%)    | ------------------------------------------------------------------------ | ------------------------------------------------------------------------ | ------------------------------------------------------------------------ |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| \- Yes                   | 3559 (79.3)                                                              | 2795 (79.2)                                                              | 764 (79.5)                                                               |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| \- No                    | 931 (20.7)                                                               | 734 (20.8)                                                               | 197 (20.5)                                                               |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| House Type, N (%)        | ------------------------------------------------------------------------ | ------------------------------------------------------------------------ | ------------------------------------------------------------------------ |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| \- Poles                 | 772 (17.2)                                                               | 613 (17.4)                                                               | 159 (16.6)                                                               |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| \- Single                | 3556 (79.2)                                                              | 2787 (79.0)                                                              | 769 (80.0)                                                               |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| \- Townhouse             | 162 (3.6)                                                                | 129 (3.6)                                                                | 33 (3.4)                                                                 |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Water Supply, N (%)      | ------------------------------------------------------------------------ | ------------------------------------------------------------------------ | ------------------------------------------------------------------------ |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| \- Pipes                 | 4051 (90.2)                                                              | 3191 (90.4)                                                              | 860 (89.5)                                                               |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| \- Other                 | 17 (0.4)                                                                 | 11 (0.3)                                                                 | 6 (0.6)                                                                  |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| \- Missing               | 422 (9.4)                                                                | 327 (9.3)                                                                | 95 (9.9)                                                                 |
+--------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+

```{r}
#create contingency table for infections and whether a house has a screen door or not
door_infection_table = table(merged_dengue_data$door_screens,
                               merged_dengue_data$any_infections)
door_infection_table

```

| Table 2        | No Infection | Any Infection |
|----------------|:------------:|:-------------:|
| Screen Door, N |     ---      |      ---      |
| \- Yes         |     623      |      192      |
| \- No          |     2906     |      769      |
