# Risk Analysis and Statistical Inference

Calculate the odds ratio for being infected with dengue by whether the house is made of concrete:

```{r}
#number of infections and no infections in concrete homes
n_infections_concrete = sum(merged_dengue_data$any_infections[merged_dengue_data$concrete_house ==1])
n_no_infections_concrete = sum(!merged_dengue_data$any_infections[merged_dengue_data$concrete_house ==1])

#calculate odds of being infected in concrete home
odds_concrete = n_infections_concrete/n_no_infections_concrete

#number of infections and no infections in non-concrete homes
n_infections_nonconcrete = sum(merged_dengue_data$any_infections[merged_dengue_data$concrete_house ==0])
n_no_infections_nonconcrete = sum(!merged_dengue_data$any_infections[merged_dengue_data$concrete_house ==0])

#calculate odds of being infected in non-concrete home
odds_nonconcrete = n_infections_nonconcrete/n_no_infections_nonconcrete

concrete_OR = odds_concrete / odds_nonconcrete
print(paste("Odds of infection by whether house is made from concrete:", concrete_OR))
```

-   The odds ratio value indicates that there is no association between the house made of concrete and being infected with dengue, although it is slightly above 1 which would mean that an individual has a slightly higher chance (1.018) of being infected with dengue if he or she does not live in a concrete home.

Run a chi-square test to assess if there is a difference in any dengue infection between males and females:

```{r}
#chi-square test
chisq.test(merged_dengue_data$sex, merged_dengue_data$any_infections)

#check which group had higher rate of infections
gender_infection_table

```

-   Although females had more infections than men in this study, since the p-value of 0.1452 is greater than the typical significance level of 0.05, we do not reject the null hypothesis, meaning that there is insufficient evidence to conclude that there is a statistically significant difference in any dengue infections between males and females.

Use a linear model to assess how the number of outdoor toilets affects the number of dengue infections. Get a 95% confidence interval for your estimate:

```{r}
#linear regression model
outdoor_toilets_n_infections_lm = lm(n_any_infections~n_toilets_outside, data = merged_dengue_data)

summary(outdoor_toilets_n_infections_lm)


# Calculate confidence intervals
confint(outdoor_toilets_n_infections_lm)
```

-   *Interpretation:* For each additional toliet, the expected change or increase of the number of dengue infections is 0.00645. However, given that the p-value is greater than 0.585 and the confidence interval includes 0, the results are not statistically significant and the number of outdoor toilets may not affect the number of dengue infections an individual or household has.

Use a logistic model to estimate the odds ratio and 95% confidence interval of any dengue infection associated with screen doors:

```{r}
screen_door_infections_glm = glm(any_infections ~ door_screens, data= merged_dengue_data, family="binomial")
summary(screen_door_infections_glm)

# Calculate confidence intervals
confint(screen_door_infections_glm)

#calculate odds ratio
exp(screen_door_infections_glm$coefficients)

```

-   *Interpretation:* The coefficient associated with having door screens is 0.15, meaning that the log odds of having a dengue infection increase by 0.15 if you do not have a screen door. The odds of having an infection without a screen door is 1.16; however, since the p-value = 0.0975, and this is greater than 0.05, this means that the odds ratio is not statistically significant.

Extend the model to estimate the odds ratio and 95% confidence interval of any dengue infection associated with screen doors, controlling for house type and garbage management:

```{r}
#logistic regression for screen doors, controlling for house type and garbage management
screen_door_infections_controlling_glm = glm(any_infections ~ door_screens + house_type + garbage_management, data= merged_dengue_data, family="binomial")
summary(screen_door_infections_glm)

# Calculate confidence intervals
confint(screen_door_infections_controlling_glm)

#calculate odds ratio
exp(screen_door_infections_controlling_glm$coefficients)
```

*Interpretation:* The odds of being infected with dengue is 1.12 if you do not have door screens, controlling for house type and garbage management; however the p-value is greater than 0.05 so it is not statistically significant.
